Deep Reinforcement Learning for Conservation Decisions

### 2021-05-27 manuscript/manuscript.Rmd
WIP

<!-- Environmental problems, ML -->
Advances in both available data and computing power have begun to open the door to a greater role for machine learning (ML) in addressing some of our planet’s most pressing environmental problems, such as the growing frequency and intensity of wildfire [@wildfire], over-exploited fisheries [@Worm2006], declining biodiversity [@biodiversity], and zoonotic pandemics [@covid].
But will ML approaches really help us tackle a changing planet?
Applications of ML in ecology have begun to illustrate the promise of two methods of ML: *supervised learning* [@joseph2020] and *unsupervised learning* [@unsupervised], but have so far largely overlooked the third and possibly most promising approach in the ML triad: *reinforcement learning* (RL).
Three features distinguish RL from other ML methods in ways that are particularly well suited to addressing global ecological change issues:

@@

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the problems considered by RL research have largely been drawn from examples in robotic movement and games like Go and Starcraft [@lillicrap; @alphazero; @alphastar]
Complex environmental problems share many similarities to these tasks and games: the need to plan many moves ahead given a large number of possible outcomes, to account for uncertainty and to respond with contingency to the unexpected.
RL agents typically develop their strategies by interacting with simulators, a practice that should not be unsettling to ecologists since learning from simulators is common across ecology.
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or climate change policy models [@dice] are already used to explore scenarios and inform ecosystem management. 
Even simpler simulators are frequently employed throughout ecology to guide understanding and response.
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of ecological models; the so called "curse of dimensionality" makes problems with a large number of states or actions intractable by conventional methods [@Possingham2006; @Marescot2013].
Neural-network-based RL techniques, referred to as *deep RL*, have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.
We believe that the application of RL methods to simulations of ecological processes will open the door to a new paradigm for tackling tough environmental change problems.
We also believe that such a transformation could bring risks and pitfalls comparable and varied as its solutions.

@@

<!-- Conceptual picture: (A) an "environment", and (B) an "agent"
  Components of an agent: 
    - *neural network*, whose *parameters* are estimated by *training* +
    - an RL algorithm, whose *hyperparameters* (including NN structure) are estimated by *tuning*
    - High-level overview of an RL algorithm components (model/model-free, on/off policy, stochastic/deterministic, etc).
  conceptual explanation of 'model-free', what it means to have a "policy" without modeling/predicting transitions

@@

An environment is a mathematical function, computer program, or real world experience that takes an agent's proposed *action* as input and returns an observation of its current *state* and an associated *reward* as output.
In contrast to classical approaches [@Marescot2013], there are few restrictions on what comprises a state or action. 
States and actions may be continuous or discrete, completely or partially observed, single or multidimensional.
The main focus of building an RL environment, however, is in the environment's transition dynamics and reward function.
The designer of the environment can make the environment follow any transition and reward function provided that both are functions of the current state and action.
This ability allows RL environments to model a broad range of decision making problems.
For example, we can set the transitions to be deterministic or stochastic.
We can also specify the reward function to be *sparse*, whereby a positive reward can only be received after a long sequence of actions, e.g. the end point in a maze.
In other environments, an agent may have to learn to forgo immediate rewards (or even accept an initial negative reward) in order to maximize the net discounted reward as we illustrate in examples here.

@@

@@

On top of the aforementioned distinctions for model-free, deep RL algorithms, there are more algorithm design choices that can be made. 
One important point of divergence is whether the agent learns a deterministic or stochastic policy.
Another is whether the algorithm is formulated to work on a discrete, continuous or both a discrete and continuous state and action space.
In table 1, we provide some of the commonly used RL algorithms and delineate some of their important characteristics.
In the appendix, we will introduce the considerations one should make in deciding whether an algorithm is well suited for a specific environment.

@@

We provide three examples that illustrate the application and potential of deep RL to ecological and conservation problems, highlighting both the potential and the inherent challenges.

@@

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(6)
txtcolor <- "#586e75"
```

@@

The first example focuses on the important but well-studied problem of setting harvest quotas in fisheries management.
Determining fishing quotas is both a critical ecological issue [@Worm2006; @Worm2009; @Costello2016], and also a textbook example that has long informed the management of renewable resources within fisheries and beyond [@Clark1990].
Given a population growth model that predicts the total biomass of a fish stock in the following year as a function of the current biomass, it is straight forward to determine what biomass corresponds to the maximum growth rate of the stock, or $B_{\textrm{MSY}}$, the biomass at Maximum Sustainable Yield (MSY) [@Schaefer1954].
When the population growth rate is stochastic, the problem is slightly harder to solve, as the harvest quota must constantly adjust to the ups and downs of stochastic growth, but it is still possible to show the optimal strategy merely seeks to maintain the stock at $B_{\textrm{MSY}}$, adjusted for any discounting of future yields [@Reed1979].

This provides a natural first benchmark for deep RL approaches, since we can compare the RL solution to the mathematical optimum directly.
The problem is also of interest because the RL approach can be extended to consider ecological dynamics that either surpass the complexity possible with SDP methods, or violate assumptions of such classical approaches).
Moreover, fisheries management is both an important global challenge in it's own right [@Worm2006; @Costello2016] as well as a common test case to understand ecological dynamics and conservation management more broadly [@something].

For illustrative purposes, we consider the simplest version of the dynamic optimal harvest problem as outlined by [@Clark1973] (for the deterministic case) and [@Reed1979] (under stochastic recruitment).
In short, the manager seeks to optimize the net present value (discounted cumulative catch) of a fishery, observing the stock size each year and setting an appropriate harvest quota in response.
In the classical approach, the best model of the fish population dynamics must first be estimated from data, potentially with posterior distributions over parameter estimates reflecting any uncertainty.
From this model, the optimal harvest policy -- that is, the a function which returns the optimal quota for each possible observed stock size -- can be determined either by analytic [@Reed1979] or numerical [@Marescot2013] methods, depending on the complexity of the model.
In a RL approach, the model is instead replaced by a computer simulation, usually referred to as an "environment." The simplest simulation is merely an implementation of the estimated model itself.
However, unlike an analytic or SDP solution that is derived directly from the estimated model, the RL approach makes no assumption about the precise functional form or parameter values underlying the dynamics -- it is in principle agnostic to the details of the simulation. <!-- ML: I think this sentence should be tightened up re RL, like one could use a model-based algorithm which uses an inputted model under the RL umbrella-->

@@

A step-by-step walk-through for training both agents on this environment is provided in the Appendix.
We compare the resulting management, policy, and reward under either RL agent to that achieved by the optimal management solution [Fig 2].
Despite having no knowledge of the underlying model and clear differences in the resulting strategies, both RL agents learn enough to achieve nearly optimal performance.

<!-- ML: I'm have some uneasiness about this figure. Is A2C tuned? I think we'd probably want to do tuning and training with a set seed so that everything is reproducible?  -->
```{r message=FALSE, include=FALSE}
sims_df <- read_csv( "figs/sims_df.csv")
policy_df <- read_csv("figs/policy_df.csv")
reward_df <- read_csv("figs/reward_df.csv")

ymin <- function(x) last(x[(ntile(x, 20)==1)])
ymax <- function(x) last(x[(ntile(x, 20)==19)])

fig_sims <- 
sims_df %>% 
  group_by(time, model) %>% 
  summarise(ymin = ymin(state),
            ymax = ymax(state),
            state = mean(state), .groups = "drop") %>%
  ggplot(aes(time, state, ymin = ymin, ymax = ymax, fill=model)) + 
  geom_ribbon(alpha= 0.3) + geom_line(aes(col = model))

fig_policy <- 
  policy_df %>% ggplot(aes(state, state - action, 
                           group=interaction(rep, model),
                           col = model)) + 
  geom_line(show.legend = FALSE) + 
  coord_cartesian(xlim = c(0, 1.3), ylim=c(0,0.9)) + 
  ylab("escapement")

```{r fig_fish, fig.cap="Fisheries management using neural network agents trained with RL algorithms A2C and TD3 compared to optimal management.  Top panel: mean fish population size over time across 100 replicates. Shaded region shows the 95% confidence interval over simulations.  Lower left: The optimal solution is policy of constant escapement. Below the target escapement of 0.5, no harvest occurs, while any stock above that level is immediately harvested back down (red line).  RL agents both adopt a policies that cease any harvest below this level, while allowing a somewhat higher escapement than optimal just above the optimal escapement level.  A2C policy is non-deterministic, and overharvest from very high stock sizes.  Lower Right: Mean cumulative reward over time across simulations. TD3 achieves nearly-optimal mean utility. Mean reward under A2C is lower, though the confidence interval still overlaps the mean reward of optimal management."}
fig_sims / ( fig_policy + fig_reward)
```

@@

Our second example focuses on a case for which we do not already have an existing, provably optimal policy to compare against.
We consider the generic problem of an ecosystem facing slowly deteriorating environmental conditions which move the dynamics closer towards a tipping point [Fig 2].
This model of a critical transition has been posited widely in ecological systems, from the simple consumer-resource model of [@May1977] on which our dynamics are based, to microbial dynamics [@Dai2012] lake ecosystem communities [@Carpenter2011] to planetary ecosystems [@Barnosky2012].
On top of these ecological dynamics we introduce an explicit ecosystem service model quantifying the value of more desirable 'high' state relative to the 'low' state.
For simplicity, we assume a proportional benefit $b$ associated with the ecosystem state $X(t)$.
Thus when the ecosystem is near the 'high' equilibrium $\hat X_H$, the corresponding ecosystem benefit $b \hat X_H$ is higher than at the low equilibrium, $b x_L$, consistent with the intuitive description of ecosystem tipping points [@Barnosky2012].
We also enumerate the possible actions which a manager may take in response to environmental degradation.
In the absence of any management response, we assume the environment deteriorates at a fixed rate $\alpha$, which can be thought of as the incremental increase in global mean temperature or similar anthropogenic forcing term.
Management can respond to slow or even reverse this trend by choosing an opposing action $A_t$.
We assume that large actions are proportionally more costly than small actions, consistent with the expectation of diminishing returns: taking the cost associated with an action $A_t$ as equal to $c A_t^2$.
Many alterations of these basic assumptions are also possible: our `gym_conservation` implements a range of different scenarios with user-configurable settings.
In each case, the manager observes the current state of the system each year and must then select the policy response that year.

@@

A close examination of the trajectories of individual simulations which cross the tipping point under either management strategy (see appendix B, fig S1, S2) further highlights the difference between these two approaches.
Under the steady-state strategy, the system remains poised too close to the tipping point: stochastic noise eventually drives most replicates across the threshold, where the steady-state strategy is too weak to bring them back once they collapse.
As replicate after replicate stochastically crashes, the mean state and mean reward bend increasingly downwards.
In contrast, the RL agent edges the system slightly farther away from the tipping point, decreasing but not eliminating the chance of a chance transition.
In the few replicates that experience a critical transition anyway, the RL agent usually responds with sufficient commitment to ensure their recovery (Fig S2).
Only 3 out of 100 replicates degrade far enough for the RL agent to give up the high cost of trying to rescue them.

Such examination makes it clear how the RL agent's use of a more dynamic strategy out-preforms the steady-state strategy.
The various kinks in the RL policy function also suggest that this solution is nevertheless surely not the optimal solution -- such features are typical of RL-based policies, where such spandral-like features from early experimentation may become baked into the long-term strategy if adjusting them has little improvement on the final reward.
Further tuning of hyper-parameters, increased training, alterations or alternatives to the training algorithm would likely be able to further improve upon this performance.
<!-- ML: I think it might be confusing to introduce RNNs, maybe just say different architectures, maybe make a footnote.  CB: agree, revised above-->

@@

<!-- Ecology has complex decisions and we lack necessary tools; RL could be it -->
Ecological challenges facing the planet today are complex and outcomes are both uncertain and consequential.
Even our best models and best research will never provide a crystal ball to the future, only better elucidate possible scenarios.
Consequently, that research must also confront the challenge of making robust, resilient decisions in a changing world.
The science of ecological management and quantitative decision-making has a long history [e.g. @Schaefer1954; @Walters1978] and remains an active area of research [@Possingham2006; @Fischer2009; @Polasky2011].
However, the limitations of classical methods such as optimal control frequently constrain applications to relatively simplified models [@Possingham2006], ignoring elements such as spatial or temporal heterogeneity and autocorrelation, stochasticity, imperfect observations, age or state structure or other sources of complexity that are both pervasive and influential on ecological dynamics [@Hastings2012].
Complexity comes not only from the ecological processes but also the available actions.
Conservation efforts can be implemented in a wide variety of ways, and the most effective policies may vary in both space and time.
While such complex dynamics and large state and action spaces are intractable to classical methods, they are straight forward to implement in simulation environments such as those used to train an RL algorithm.
Neural networks used in deep RL have proven remarkably effective in handling such complexity, particularly when leveraging immense computing resources increasingly available through advances in hardware and software [@gpu_computing].

<!-- Model free is great bc we don't know the true models -->
The rapidly expanding class of Model-free RL algorithms is particularly appealing given the ubiquitous presence of model uncertainty in ecological dynamics.
Rarely do we know the underlying functional forms for ecological processes -- is population recruitment more Beverton-Holt or Ricker?
Methods which must first assume something about the structure or functional form of a process before estimating the corresponding parameter can only ever be as good as those structural assumptions.
Frequently available ecological data is insufficient to distinguish between possible alternative models [@devalpine], or the correct model may be non-identifiable with any amount of data.
Model-free RL approaches offer a powerful way to factor this thorny issue.
The model-free RL agent needs to make no such a priori hypothesis about a particular functional form.
Model-free approaches have proven successful at learning effective policies even when the underlying model may be difficult or impossible to learn [@Pong2020], as long as simulations of possible mechanisms are available.

<!-- Real problems are harder than these examples -->
<!-- ML: I feel like this paragraph could be condensed with the one above -->
<!-- CB: I think "model free is great" and "real problems are harder than these examples" are two distinct messages though and at least one would get lost in mashing them together.  Perhaps the message doesn't come across in the current version.  Does look like this paragraph is similar in overall message to the next one though, maybe could be combined with that?  -->
The examples presented here only scrape the surface of possible RL applications to conservation problems.
The examples we have focused on are intentionally quite simple, though it is worth remembering that these very same simple models have a long history of relevance and application in both research and policy contexts.
Despite their simplicity, the optimal strategy is not always obvious before hand, however intuitive it may appear in retrospect.
In the case of the ecosystem tipping point scenario, the optimal strategy is unknown, and the approximate solution found by our RL implementation could almost certainly be improved upon.
Even at it's best, an RL-trained agent can only be as good as the scenarios on which it has been trained.
In these simple examples in which the simulation implements a single model, training is analogous to classical methods which take the model as given [@Marescot2013].
But classical approaches can be difficult to generalize when the underlying model is unknown: alternative models may require a completely different approach.
In contrast, the process of training an RL algorithm on a more complex problem is no different than the simple one: we only need access to a simulation which can generate plausible future states in response to possible actions.
Many of our most realistic ecological models exist only as numerical simulations of lower-level dynamics, such as the individual-based models used in the management of forests and wildfire [@sortie; @wildfire], disease [@covid], marine ecosystems [@ecopath], or global climate change [@dice].

<!-- Solving 'real' problems will be hard -->
Successfully applying RL to complex ecological problems is no easy task.
The flexibility of the approach makes it relatively straightforward to include realistic biology and environmental science in framing an RL problem, but training an RL agent to reach an acceptable level of performance can often be more challenging than expected.
As the examples we have presented illustrate, it is important to begin with simple problems, including those for which an optimal strategy is already known.
Such examples provide important benchmarks against which to calibrate the performance, tuning and training requirements of RL,
and can easily be extended into more complex problems once RL agents can master the basics.
Finding acceptable policies even in relatively simple problems may prove beyond the reach of existing algorithms and available computing power.
In seeking either to adopt or assess a deep RL-based approach to an ecological issue, we highlight three key issues that are important to bear in mind: generalizability, interpretability, and reward design.

<!-- reason 1: generalizability-->
<!-- ML: pivoting this to generalizability but still rewrote past paragraph which might fit in the appendix somewhere-->
Currently, deep RL agents struggle to generalize their skills [@generalization].
This inability can affect performance during training and at evaluation. 
For instance, training an agent on an environment that demands learning an assortment of skills is typically much more difficult than training an agent to learn one specific skill [@CITE].
And, at evaluation, if you give an agent a novel task, including tasks that are closely related to the examples that the agent saw in training, the agent can easily fail to generalize its learned behavior to this new situation [@CITE].
These failures are all due to the brittleness of deep neural networks. 
Across deep learning, neural networks tend to perform exceptionally well on narrowly defined problems, yet neural networks show great fragility when given new situations or problem statements that require flexibility [@CITE].
The ability to adapt to new situations is critical for any decision-making problem in the real world, so the current lack of this ability in deep RL may limit deep RL's effectiveness to conservation problems.

>>>>

<!-- Environmental problems, ML -->
Advances in both available data and computing power have begun to open the door to a greater role for machine learning (ML) in addressing some of our planet’s most pressing environmental problems, such as the growing frequency and intensity of wildfire [@wildfire], over-exploited fisheries [@worm2006], declining biodiversity [@biodiversity], and zoonotic pandemics [@covid].
But will ML approaches really help us tackle a changing planet?
Applications of ML in ecology have begun to illustrate the promise of two methods of ML: *supervised learning* [@joseph2020] and *unsupervised learning* [@unsupervised], but have so far largely overlooked the third and possibly most promising approach in the ML triad: *reinforcement learning* (RL).
Three features distinguish RL from other ML methods in ways that are particularly well suited to addressing global ecological change issues:

@@

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the problems considered by RL research have largely been drawn from examples in robotic movement and games like Go and Starcraft [@lillicrap; @alphazero; @alphastar]
Complex environmental problems share many similarities to these tasks and games: the need to plan many moves ahead given a large number of possible outcomes, to account for uncertainty and to respond with contingency to the unexpected.
RL agents typically develop their strategies by interacting with simulators, a practice that should not be unsettling to ecologists since learning from simulators is common across ecology.
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or global circulation models (GCMs) in climate [@gcm] are already used to explore scenarios and inform ecosystem management. 
Even simpler simulators are frequently employed throughout ecology to guide understanding and response.
<!-- Kari: ick, jargon heavy.  Need to say: some existing methods, but limited to simple problems or simple solutions -->
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of ecological models; the so called "curse of dimensionality" makes problems with a large number of states or actions intractable by conventional methods [@Possingham2006; @Marescot2013].
Neural-network-based RL techniques, referred to as *deep RL*, have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.
We believe that the application of RL methods to simulations of ecological processes will open the door to a new paradigm for tackling tough environmental change problems.
We also believe that such a transformation could bring risks and pitfalls comparable and varied as its solutions.

@@

<!-- Conceptual picture: (A) an "environment", and (B) an "agent"

What do we mean by an "environment"?  
  - Transition function
  - "reward"
  - examples

What do we mean by an "agent"?

  Model-based vs Model free.  We focus on Model free agents which use Neural networks, a subset of deep RL

  Components of an agent: 
    - *neural network*, whose *parameters* are estimated by *training* +
    - an RL algorithm, whose *hyperparameters* (including NN structure) are estimated by *tuning*
    - High-level overview of an RL algorithm components (model/model-free, on/off policy, stochastic/deterministic, etc).

  conceptual explanation of 'model-free', what it means to have a "policy" without modeling/predicting transitions

@@

An environment is a mathematical function, computer program, or real world experience that takes an agent's proposed *action* as input and returns an observation of its current *state* and an associated *reward* as output.
<!-- ML: what are classical approaches below here? -->
In contrast to classical approaches, there are few restrictions on what comprises a state or action. 
The state may be continuous or discrete, completely or partially observed, single or multidimensional.
For instance, we might represent an ecological community of 10 interacting species with a 10-dimensional continuously valued state space signifying the population density of each species.
The action space can be similarly complex, since an action can be continuous or discrete, single or multidimensional.
For the previous environment, we might set the action space to be continuously valued and 10-dimensional, where each dimension corresponds to the amount of conservation effort aimed at the different species over a time step.   
The main focus of building an RL environment, however, is in the environment's transition dynamics and reward function.
The designer of the environment can make the environment follow any transition and reward function provided that both are functions of the current state and action.
This ability allows RL environments to model a broad range of decision making problems.
For example, we can set the transitions to be deterministic or stochastic.
We can also specify the reward function to be *sparse*, whereby a positive reward can only be received after a long sequence of actions, e.g. the end point in a maze.
In other environments, an agent may have to learn to forgo immediate rewards (or even accept an initial negative reward) in order to maximize the net discounted reward as we illustrate in examples here.

@@

**Survey of RL Algorithms**

@@

On top of the aforementioned distinctions for model-free, deep RL algorithms, there are more algorithm design choices that can be made. 
One important point of divergence is whether the agent learns a deterministic or stochastic policy.
Another is whether the algorithm is formulated to work on a discrete, continuous or both a discrete and continuous state and action space.
In table **, we provide some of the commonly used RL algorithms and delineate some of their important characteristics.
In the appendix, we will introduce the considerations one should make in deciding whether an algorithm is well suited for a specific environment.

<!-- Kari: I think we need more background on what a neural network is, how it fits in, and what other options are - maybe use examples of hyperparameters that would be in all RL approaches not just those that use neural nets -->

<!-- Can we add citations to each please -->
<!--

Should we mention:

- Temporal Differencing, RIENFORCE
- SARSA, 
- Double DQN?  
Can we add HER (since delayed reward case may be particularly interesting)
Should we do anything to better reflect the 'family tree' here? (e.g. DQN -> DDPG -> TD3)

@@

We provide three examples that illustrate the application and potential of deep RL to ecological and conservation problems, highlighting both the potential and challenges there-in.

@@

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(6)
txtcolor <- "#586e75"

source("../R/plotting-helpers.R")
```

@@

The first example focuses on the important but well-studied problem of setting harvest quotas in fisheries management.
Determining fishing quotas is both a critical ecological issue [@Worm2006; @Worm2009; @Costello2018], and also a textbook example that has long informed the management of renewable resources within fisheries and beyond [@Clark1990].
Given a population growth model that predicts the total biomass of a fish stock in the following year as a function of the current biomass, it is straight forward to determine what biomass corresponds to the maximum growth rate of the stock, or $B_{\textrm{MSY}}$, the biomass at Maximum Sustainable Yield (MSY) [@Schaefer1954].
When the population growth rate is stochastic, the problem is slightly harder to solve, as the harvest quota must constantly adjust to the ups and downs of stochastic growth, but it is still possible to show the optimal strategy merely seeks to maintain the stock at $B_{\textrm{MSY}}$, adjusted for any discounting of future yields [@Reed1979].

This provides a natural first benchmark for deep RL approaches, since we can compare the RL solution to the mathematical optimum directly.
The problem is also of interest because the RL approach can be extended to consider ecological dynamics that either surpass the complexity possible with SDP methods, or violate assumptions of such classical approaches).
Moreover, fisheries management is both an important global challenge in it's own right [@Worm2006; @Costello2018] as well as a common test case to understand ecological dynamics and conservation management more broadly [@something].

For illustrative purposes, we consider the simplest version of the dynamic optimal harvest problem as outlined by [@Clark1974] (for the deterministic case) and [@Reed1979] (under stochastic recruitment).
In short, the manager seeks to optimize the net present value (discounted cumulative catch) of a fishery, observing the stock size each year and setting an appropriate harvest quota in response.
In the classical approach, the best model of the fish population dynamics must first be estimated from data, potentially with posterior distributions over parameter estimates reflecting any uncertainty.
From this model, the optimal harvest policy -- that is, the a function which returns the optimal quota for each possible observed stock size -- can be determined either by analytic [@Reed1979] or numerical [@Marescot2013] methods, depending on the complexity of the model.
In a RL approach, the model is instead replaced by a computer simulation, usually referred to as an "environment." The simplest simulation is merely an implementation of the estimated model itself.
However, unlike an analytic or SDP solution that is derived directly from the estimated model, the RL approach makes no assumption about the precise functional form or parameter values underlying the dynamics -- it is in principle agnostic to the details of the simulation. <!-- ML: I think this sentence should be tightened up re RL, like one could use a model-based algorithm which uses an inputted model under the RL umbrella-->

@@

A step-by-step walk-through for training both agents on this environment is provided in the Appendix.
We compare the resulting management, policy, and reward under either RL agent to that achieved by the optimal management solution [Fig 1].
Despite having no knowledge of the underlying model and clear differences in the resulting strategies, both RL agents learn enough to achieve nearly optimal performance.

```{r message=FALSE, include=FALSE}
sims_df <- read_csv( "figs/sims_df.csv") %>% filter(model %in% c("Optimal", "TD3_tuned"))
policy_df <- read_csv("figs/policy_df.csv") %>% filter(model %in% c("Optimal", "TD3_tuned"))

```{r fig_fish, fig.cap="Fisheries management using neural network agents trained with RL algorithms A2C and TD3 compared to optimal management.  Top panel: mean fish population size over time across 100 replicates. Shaded region shows the 95% confidence interval over simulations.  Lower left: The optimal solution is policy of constant escapement. Below the target escapement of 0.5, no harvest occurs, while any stock above that level is immediately harvested back down (red line).  RL agents both adopt a policies that cease any harvest below this level, while allowing a somewhat higher escapement than optimal just above the optimal escapement level.  A2C policy is non-deterministic, and overharvest from very high stock sizes.  Lower Right: Mean cumulative reward over time across simulations. TD3 achieves nearly-optimal mean utility. Mean reward under A2C is lower, though the confidence interval still overlaps the mean reward of optimal management."}
plot_sims(sims_df) / ( plot_policy(policy_df) + plot_reward(sims_df))
```

@@

Our second example focuses on a case for which we do not already have an existing, provably optimal policy to compare against.
We consider the generic problem of an ecosystem facing slowly deteriorating environmental conditions which move the dynamics closer towards a tipping point [Fig 2].
This model of a critical transition has been posited widely in ecological systems, from the simple consumer-resource model of [@May1977] on which our dynamics are based, to microbial dynamics [@Dai2013] lake ecosystem communities [@Carpenter2008] to planetary ecosystems [@Barnkosky2014].
On top of these ecological dynamics we introduce an explicit ecosystem service model quantifying the value of more desirable 'high' state relative to the 'low' state.
For simplicity, we assume a proportional benefit $b$ associated with the ecosystem state $X(t)$.
Thus when the ecosystem is near the 'high' equilibrium $\hat X_H$, the corresponding ecosystem benefit $b \hat X_H$ is higher than at the low equilibrium, $b x_L$, consistent with the intuitive description of ecosystem tipping points [@Barnosky2014].
We also enumerate the possible actions which a manager may take in response to environmental degradation.
In the absence of any management response, we assume the environment deteriorates at a fixed rate $\alpha$, which can be thought of as the incremental increase in global mean temperature or similar anthropogenic forcing term.
Management can respond to slow or even reverse this trend by choosing an opposing action $A_t$.
We assume that large actions are proportionally more costly than small actions, consistent with the expectation of diminishing returns: taking the cost associated with an action $A_t$ as equal to $c A_t^2$.
Many alterations of these basic assumptions are also possible: our `gym_conservation` implements a range of different scenarios with user-configurable settings.
In each case, the manager observes the current state of the system each year and must then select the policy response that year.

@@

A close examination of the trajectories of individual simulations which cross the tipping point under either management strategy (see appendix, fig S1, S2) further highlights the difference between these two approaches.
Under the steady-state strategy, the system remains poised too close to the tipping point: stochastic noise eventually drives most replicates across the threshold, where the steady-state strategy is too weak to bring them back once they collapse.
As replicate after replicate stochastically crashes, the mean state and mean reward bend increasingly downwards.
In contrast, the RL agent edges the system slightly farther away from the tipping point, decreasing but not eliminating the chance of a chance transition.
In the few replicates that experience a critical transition anyway, the RL agent usually responds with sufficient commitment to ensure their recovery (Fig S2).
Only 3 out of 100 replicates degrade far enough for the RL agent to give up the high cost of trying to rescue them.

Such examination makes it clear how the RL agent's use of a more dynamic strategy out-preforms the steady-state strategy.
The various kinks in the RL policy function also suggest that this solution is nevertheless surely not the optimal solution -- such features are typical of RL-based policies, where such spandral-like features from early experimentation may become baked into the long-term strategy if adjusting them has little improvement on the final reward.
Further tuning of hyper-parameters, increased training, alterations to the training algorithm or an entirely different training algorithm would likely be able to further improve upon this performance.
The use of a recurrent neural network (RNN) architecture instead of the multi-layer perceptron (MLP) used here would open the door for the policy to reflect the non-autonomous dynamics more directly. 
<!-- ML: I think it might be confusing to introduce RNNs, maybe just say different architectures, maybe make a footnote-->

@@

<!-- Ecology has complex decisions and we lack necessary tools; RL could be it -->
Ecological challenges facing the planet today are complex and outcomes are both uncertain and consequential.
Even our best models and best research will never provide a crystal ball to the future, only better elucidate possible scenarios.
Consequently, that research must also confront the challenge of making robust, resilient decisions in a changing world.
The science of ecological management and quantitative decision-making has a long history [e.g. @Schaefer1954; @Walters1974] and remains an active area of research [@Possingham2006; @Fischer2009; @Polasky2011].
However, the limitations of classical methods such as optimal control frequently constrain applications to relatively simplified models [@Possingham2006], ignoring elements such as spatial or temporal heterogeneity and autocorrelation, stochasticity, imperfect observations, age or state structure or other sources of complexity that are both pervasive and influential on ecological dynamics [@HastingsEncylopedia].
Complexity comes not only from the ecological processes but also the available actions.
Conservation efforts can be implemented in a wide variety of ways, and the most effective policies may vary in both space and time.
While such complex dynamics and large state and action spaces are intractable to classical methods, they are straight forward to implement in simulation environments such as those used to train an RL algorithm.
Neural networks used in deep RL have proven remarkably effective in handling such complexity, particularly when leveraging immense computing resources increasingly available through advances in hardware and software [@gpu_computing].

<!-- Model free is great bc we don't know the true models -->
The rapidly expanding class of Model-free RL algorithms is particularly appealing given the ubiquitous presence of model uncertainty in ecological dynamics.
Rarely do we know the underlying functional forms for ecological processes -- is population recruitment more Beverton-Holt or Ricker?
Methods which must first assume something about the structure or functional form of a process before estimating the corresponding parameter can only ever be as good as those structural assumptions.
Frequently available ecological data is insufficient to distinguish between possible alternative models [@devalpine], or the correct model may be non-identifiable with any amount of data.
Model-free RL approaches offer a powerful way to factor this thorny issue.
The model-free RL agent needs to make no such a priori hypothesis about a particular functional form. <!-- ML: I feel like the remaining part of this dont full drive home the benefit of model-free rl, like its able to learn good performing policies even in the case when a model is very difficult or impossible to learn. -->
Alternative models can easily be implemented in simulations, allowing an agent to train against both contingencies.
A successful agent will learn a policy that works well in both scenarios.
This is analogous to classical approaches for integrating over parameter or model uncertainty [e.g. @Walters1981; @Ludwig1982],

<!-- Real problems are harder than these examples -->
<!-- ML: I feel like this paragraph could be condensed with the one above -->
The examples presented here only scrape the surface of possible RL applications to conservation problems.
The examples we have focused on are intentionally quite simple, though it is worth remembering that these very same simple models have a long history of relevance and application in both research and policy contexts.
Despite their simplicity, the optimal strategy is not always obvious before hand, however intuitive it may appear in retrospect.
In the case of the ecosystem tipping point scenario, the optimal strategy is unknown, and the approximate solution found by our RL implementation could almost certainly be improved upon.
Even at it's best, an RL-trained agent can only be as good as the scenarios on which it has been trained.
In these simple examples in which the simulation implements a single model, training is analogous to classical methods which take the model as given [@Marescot2013].
But classical approaches can be difficult to generalize when the underlying model is unknown: alternative models may require a completely different approach.
In contrast, the process of training an RL algorithm on a more complex problem is no different than the simple one: we only need access to a simulation which can generate plausible future states in response to possible actions.
Many of our most realistic ecological models exist only as numerical simulations of lower-level dynamics, such as the individual-based models used in the management of forests and wildfire [@sortie; @firemodel], disease [@disease], marine ecosystems [@ecopath], or global climate change [@gcm].

<!-- Solving 'real' problems will be hard -->
Successfully applying RL to complex ecological problems is no easy task.
The flexibility of the approach makes it relatively straightforward to include realistic biology and environmental science in framing an RL problem, but training an RL agent to reach an acceptable level of performance is another task.
As the examples we have presented illustrate, it is important to begin with simple problems, including those for which an optimal strategy is already known.
Such examples provide important benchmarks against which to calibrate the performance, tuning and training requirements of RL,
and can easily be extended into more complex problems once RL agents can master the basics.
Finding acceptable policies even in relatively simple problems may prove beyond the reach of existing algorithms and available computing power.
In seeking either to adopt or assess a deep RL-based approach to an ecological issue, we highlight three key issues that are important to bear in mind: instability, interpretability, and reward design.

<!-- reason 1: instability-->
<!-- ML: I think this needs to be tightened up a bit-->
One leading challenge for deep RL today is the problem of instability.
RL agents trained on narrowly defined environments may perform poorly even when the environment is altered in seemingly trivial way, such as a re-scaling of parameters.
To avoid this, researchers should seek to both train and evaluate agents against a range of environments representing different parameters or processes dynamics.
Stable performance can be challenging even when using the same environment.
For a range of reasons, including high variance [@cite] and the exploration vs. exploitation trade-off [@cite], model-free and model-based algorithms tend to have unstable performance on environments with non-tabular state-action spaces [cite RL that matters].
For instance, it is not uncommon to observe vastly different behavior when an agent is initialized with different random seeds [cite ibid].
Such unstable behavior can often be avoided by alternate selection of hyperparameters, so researchers should consider stability as well as performance during hyperparameter selection.
The choice of RL algorithm, design of the reward function, and duration of training can also influence stability.

<!--
Under some conditions, RL agents can prove adept at _transfer learning_, in which an agent trained in one environment performs well under a different environment.
-->
<!-- if we're really recommending 'testing' for stability, maybe we need to demo that in the appendix-->

<!-- reason 2: interpretability -->
<!-- ML: Re our OU talk, might be worth saying that this is a barrier to adoption? If we want to ultimately use these RL agents to make important decisions, it would be hard to swallow us blindly using them without having any sense to what is going on. -->
Interpretability is a common challenge at the adoption of most ML methods [@cite]. 
Although there have been some attempts to increase the transparency of RL agents that use neural network as function approximators [cite], it is still very difficult to explain the behavior of RL-trained agents.
As our examples have illustrated, it may be possible to gain intuition on an agent's behavior by visualizing the action selected against each possible state, though this becomes more difficult in large problems.
Another challenge is that Deep RL algorithms are very sample inefficient, all though progress is being made to improve this [@cite]. <!-- ML: I think we can cut this sentence, mentionned it in passing earlier-->
-->
